---
title: "In-season estimates of ad-intact fish returning to the Krassel PIT-array"
author:
- name: Mike Ackerman
  affiliation: Research Scientist
- name: Ryan N. Kinzer
  affiliation: Fisheries Data Analyst and Coordinator
- name: Jason Vogel
  affiliation: Research Division Director
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  bookdown::pdf_document2:
    citation_package: biblatex #natbib
    number_sections: no
    keep_tex: yes
    fig_caption: yes
    latex_engine: pdflatex #xelatex
    template: NPT_report_template2.tex
  bookdown::html_document2:
    number_sections: no
    fig_caption: yes
    toc: yes
    floating_toc: yes
    df_print: paged
shorttitle: Ad-Intact Estimates
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
branding: yes
bibliography:
- references.bib
link-citations: true
csl: "american-fisheries-society.csl" # Insert path for the bib-style
---

```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
library(tidyverse)

opts_chunk$set(echo = FALSE, message = FALSE, error = FALSE,
                      warning = FALSE)

options(knitr.kable.NA = '-')
```


## Summary 
* Current PIT-tag expansions at the Krassel PIT-array (KRS; as of July 17th)^[Scripts to reproduce these summaries are available at "https://github.com/NPTfisheries/sf_salmon_harvest.com".]:
  * 656 natural-origin fish estimated from 118 observed tags and 18% sampling/tagging rate (Tables 1 and 2)
  * Using a 20% tagging rate the estimate is 590.
* KRS was non-operational: 7/10, 7/11, 7/14, 7/15, and 7/16
  * Natural-origin PIT-tag passage during this period is estimated as $\le$ 10 tags for an additional 56 natural-origin fish.
* The average passage date of $\ge$ 75% of the total return across KRS is July 15th (Figure \@ref(fig:run-timing)).
* End of season natural-origin estimate: 948
  * NOR: (656 + 56)/.75 = 949
* Current Lower Granite Estimate of McCall Integrated fish is 465.
  * Estimated at KRS after accounting for SbyC is 443 (95% conversion)
  
* **Total ad-intact fish at KRS: 1,413**
  * 949 (NOR) + 465 (Integrated) = 1,414
  
## Assumptions
* 100% conversion of integrated fish to KRS.
* 100% detection efficiency at KRS
  * In 2021 and 2022 KRS detections efficiency was $\ge$ 95%.
* Tagging rate at LGR of 18%.
  * Only 4 out of 118 NOR tags were marked at LGR during the 29% sampling rate, all others marked at 18%.
* Static interpolation rate of 2 tags during the missed period at KRS when observations are typically decreasing.

```{r process-data}
load("../data/sfsr_obs.rda")

sfsr_fish <- sfsr_obs %>%
  filter(mark_site %in% c('LGRLDR', 'MCCA')) %>%
  filter(rel_site %in% c('LGRLDR', 'KNOXB')) %>%
  mutate(julian = lubridate::mdy(paste0(format(min_det, format = '%m-%d'), '-2020'))) %>%
  mutate(group = case_when(
    mark_site == 'MCCA' & grepl('AD', flags) ~ 'Segregated',
    mark_site == 'MCCA' & grepl('CW||AI', flags) ~ 'Integregated',
    mark_site == 'LGRLDR' & mark_rear_type_name != 'H' ~ 'Natural',
    TRUE ~ NA
  ))
  
obs <- sfsr_fish %>%
  filter(group != 'other') %>%
  filter(site_code %in% c('SFG', 'KRS')) %>%
  group_by(site_code, tag_code) %>%
  slice_min(min_det)

# check run-timing by group - not a big difference
obs_sum <- obs %>%
  group_by(site_code, spawn_year, julian) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  complete(site_code, spawn_year, julian, fill = list(n=0)) %>%
  group_by(site_code, spawn_year) %>%
  mutate(csum = cumsum(n),
         cdf = csum/sum(n)) %>%
  ungroup()

avg_run <- obs_sum %>%
  filter(spawn_year != 2023) %>%
  group_by(site_code, julian) %>%
  summarise(cdf = mean(cdf))

```


```{r load-data}
load('../data/expansion.rda') 
```


```{r exp-table}
library(flextable)

tag_exp %>%
  filter(site_code == 'KRS') %>%
  mutate(rel_group = gsub('McCall - |LGR - ','',rel_group),
         rel_year = as.character(rel_year)) %>%
  select(rel_year, rel_group, SbyC, tag_expansion, n_tags, est) %>%
  arrange(rel_year, rel_group) %>%
  flextable() %>%
  set_header_labels(rel_year = 'Release Year', rel_group = 'Group', SbyC = 'SbyC', tag_expansion = 'Expansion', n_tags = 'n_tags', est = 'Estimate') %>%
  set_caption(
    caption = 'Current tag expansion rates and estimates for SFSR release groups at KRS.',
    style = "Table Caption",
    autonum = officer::run_autonum(seq_id = "tab", bkm = "tab1")) %>%
  theme_booktabs() %>%
  set_table_properties(align = "center", layout = "fixed")
```

```{r exp-table-sum}
exp_df %>%
  filter(site_code == 'KRS') %>%
  mutate(rel_group = gsub('McCall - |LGR - ','',rel_group)) %>%
  select(rel_group, n_tags, estimate = n_tags_exp) %>%
  arrange(rel_group) %>%
  flextable() %>%
  set_header_labels(rel_group = 'Group', n_tags = 'n_tags', estimate = 'Estimate') %>%
  set_caption(
    caption = 'Current SFSR release group estimates at KRS.',
    style = "Table Caption",
    autonum = officer::run_autonum(seq_id = "tab", bkm = "tab1")) %>%
  theme_booktabs() %>%
  set_table_properties(align = "center", layout = "fixed")
```


```{r run-timing, fig.width=6, fig.height=5, fig.cap = "Run-timing of SFSR hatchery (segregated and integrated) and natural origin returns across PIT-arrays. The grey lines indicate individual spawn year returns for 2010-2022, the dark blue line is the average cumulative proportion of returns across the same years, and the green line represents the current 2023 return."}
ggplot(data = NULL, aes(x = julian, y = cdf)) +
    geom_line(data = obs_sum %>% filter(spawn_year != 2023), aes(group = spawn_year), colour = 'grey') +
    geom_line(data = avg_run, colour = 'navy') +
    geom_line(data = obs_sum %>% filter(spawn_year == 2023), colour = 'limegreen') +
    scale_x_date(date_breaks = '2 weeks', date_labels = "%b-%d") +
    facet_wrap(~fct_rev(site_code), ncol = 1) +
  theme_bw() +
  labs(x = 'Date',
       y = 'CDF')
```
