---
title: "In-season estimates of ad-intact fish returning to the Krassel PIT-array"
author:
- name: Mike Ackerman
  affiliation: Research Scientist
- name: Ryan N. Kinzer
  affiliation: Fisheries Data Analyst and Coordinator
- name: Jason Vogel
  affiliation: Research Division Director
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  bookdown::pdf_document2:
    citation_package: biblatex #natbib
    number_sections: no
    keep_tex: yes
    fig_caption: yes
    latex_engine: pdflatex #xelatex
    template: NPT_report_template2.tex
  bookdown::html_document2:
    number_sections: no
    fig_caption: yes
    toc: yes
    floating_toc: yes
    df_print: paged
shorttitle: Ad-Intact Estimates
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
branding: yes
bibliography:
- references.bib
link-citations: true
csl: "american-fisheries-society.csl" # Insert path for the bib-style
---

```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
library(tidyverse)

opts_chunk$set(echo = FALSE, message = FALSE, error = FALSE,
                      warning = FALSE)

options(knitr.kable.NA = '-')
```

```{r load-data}
load('../data/expansion.rda') 
```

## Summary 
* Current PIT-tag expansions at the Krassel PIT-array (KRS; as of July 17th)^[Scripts to reproduce these summaries are available at "https://github.com/NPTfisheries/sf_salmon_harvest.com".]:
  * 655 natural-origin fish estimated from 118 observed tags
  * 560 integrated fish estimated from 80 tags with a ratio of approximately 7 fish/tag
* KRS was non-operational: 7/10, 7/11, 7/14, 7/15, and 7/16
  * Integrated and natural-origin PIT-tag passage during this period is estimated as $\le$ 10 for each group.
  * additional 70 integrated fish
  * additional 56 natural-origin
* Ad-intact fish currently estimated above KRS: 1,341
* The average passage date of $\ge$ 75% of the total return across KRS is July 15th (Figure \@ref(fig:run-timing)).
* End of season ad-intact total estimate: 1,788
  * NOR: (655 + 56)/.75 = 948
  * Integrated: (560 + 70)/.75 = 840

## Assumptions
* 100% detection efficiency at KRS
  * In 2021 and 2022 KRS detections efficiency was $>$95%.
* Static interpolation rate of 2 tags when observations are typically decreasing.
* Integrated fish observed at KRS is approximately 100 fish higher than the estimate at LGR.
* Integrated tag ratio of 7 is slighter higher than actual rate for each release group.

```{r process-data}
load("../data/sfsr_obs.rda")

sfsr_fish <- sfsr_obs %>%
  filter(mark_site %in% c('LGRLDR', 'MCCA')) %>%
  filter(rel_site %in% c('LGRLDR', 'KNOXB')) %>%
  mutate(julian = lubridate::mdy(paste0(format(min_det, format = '%m-%d'), '-2020'))) %>%
  mutate(group = case_when(
    mark_site == 'MCCA' & grepl('AD', flags) ~ 'Segregated',
    mark_site == 'MCCA' & grepl('CW||AI', flags) ~ 'Integregated',
    mark_site == 'LGRLDR' & mark_rear_type_name != 'H' ~ 'Natural',
    TRUE ~ NA
  ))
  
obs <- sfsr_fish %>%
  filter(group != 'other') %>%
  filter(site_code %in% c('SFG', 'KRS')) %>%
  group_by(site_code, tag_code) %>%
  slice_min(min_det)

# check run-timing by group - not a big difference
obs_sum <- obs %>%
  group_by(site_code, spawn_year, julian) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  complete(site_code, spawn_year, julian, fill = list(n=0)) %>%
  group_by(site_code, spawn_year) %>%
  mutate(csum = cumsum(n),
         cdf = csum/sum(n)) %>%
  ungroup()

avg_run <- obs_sum %>%
  filter(spawn_year != 2023) %>%
  group_by(site_code, julian) %>%
  summarise(cdf = mean(cdf))
  

```


```{r run-timing, fig.width=6, fig.height=5, fig.cap = "Run-timing of SFSR hatchery (segregated and integrated) and natural origin returns across PIT-arrays. The grey lines indicate individual spawn year returns for 2010-2022, the dark blue line is the average cumulative proportion of returns across the same years, and the green line represents the current 2023 return."}
ggplot(data = NULL, aes(x = julian, y = cdf)) +
    geom_line(data = obs_sum %>% filter(spawn_year != 2023), aes(group = spawn_year), colour = 'grey') +
    geom_line(data = avg_run, colour = 'navy') +
    geom_line(data = obs_sum %>% filter(spawn_year == 2023), colour = 'limegreen') +
    scale_x_date(date_breaks = '2 weeks', date_labels = "%b-%d") +
    facet_wrap(~fct_rev(site_code), ncol = 1) +
  theme_bw() +
  labs(x = 'Date',
       y = 'CDF')
```
