---
title: "In-season Escapement Estimates to SFG and KRS PIT-tag Arrays in the South Fork Salmon River"
author:
  - name: "Mike Ackerman"
    email: mikea@nezperce.org
    affiliation: "Research Scientist"
    correspondence: yes

  - name: "Ryan Kinzer"
    affiliation: "Fisheries Data Analysis Coordinator"

  - name: "Jason Vogel"
    affiliation: "Research Division Director"

date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    theme: cerulean
---

<!-- the following inserts the NPT logo into header -->
```{=html}
<script>
   $(document).ready(function() {
     $head = $('#header');
     $head.prepend('<img src=\"DFRM.png\" style=\"float: right;width: 150px;\"/>')
   });
</script>

<style>
p.caption {
  font-size: 0.8em;
}
</style>
```

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE
)
```

```{r load-libraries}
# for formatting
library(kableExtra)

# load necessary libraries
library(tidyverse)
library(lubridate)
library(readxl)
library(here)
library(janitor)
library(PITcleanr)

```

```{r load-obs-data}
# load latest compiled observation dataset based on dttm
load(here("data/sfsr_chnk_compiled_obs_latest.rda"))
current_year = year(update_dttm)

# filter to current_year
sfsr_comp_obs_df_yr = sfsr_comp_obs_df %>%
  filter(spawn_yr == current_year)

```

```{r set-parameters}
# lgr trap and subsample rates
lgr_trap_rate         = 0.186       # assumed lgr trapping/tagging rate
#lgr_hor_chnk_sub_rate = 0.33333333  # assumed lgr hatchery chinook subsample rate

# chnk size cutoffs
jack_cutoff_mm        = 580
adult_2salt_cutoff_mm = 815

# paths to mark/expansion rates
lgr_mark_filepath     = here("data/mark_rates/2026 PIT_Tag_Analysis_LowerGranite.xlsx") # path to lgr tag-specific expansion data
bon_exp_filepath      = here("data/mark_rates/2026 PIT_Tag_Analysis_Bonneville.xlsx")   # path to past bon expansion rates (backup option)

# array sets
nodes_all      = c("SFG", "KRS", "KRS_D", "KRS_U", "STR", "ESS_D", "ESS_U", "ZEN_D", "ZEN_U")
nodes_report   = c("SFG", "KRS_D", "ESS_D", "ZEN_D")
nodes_upstream = c("KRS_U", "ESS_U", "ZEN_U")

```

```{r assign-rg-helper}
assign_rel_group = function(rel_site,
                            flags,
                            mark_rear_type_name) {
  case_when(
    rel_site == "KNOXB"  & str_detect(flags, "AI")    ~ "McCall - Integrated",
    rel_site == "KNOXB"  & str_detect(flags, "AD")    ~ "McCall - Segregated",
    rel_site == "LGRLDR" & mark_rear_type_name == "W" ~ "LGRLDR",
    rel_site == "LGRLDR" & str_detect(flags, "CW")    ~ "LGRLDR",
    TRUE ~ NA_character_
  )
}

```

```{r load-exp-data}
# load lgr tag expansions, the preferred option, for current year
lgr_tag_exp = read_excel(path = lgr_mark_filepath,
                         sheet = "PIT Data") %>%
  clean_names() %>%
  filter(!is.na(tag)) %>%
  select(
    tag,
    brood_year,
    release_site_code,
    ocean_age,
    sby_c,
    exp_rate = expansion
  )

# load historic, summarized mark rates from bonneville, the backup option
bon_exp_rates = read_excel(path = bon_exp_filepath,
                           sheet = "Historic Juv Rel Numbers") %>%
  clean_names() %>%
  filter(str_detect(hatchery, "McCall")) %>%
  mutate(
    hatchery = case_when(
      hatchery == "McCall (Int)" ~ "McCall - Integrated",
      hatchery == "McCall (Seg)" ~ "McCall - Segregated",
      TRUE ~ hatchery
    )
  ) %>%
  mutate(
    ral_mark_rate = pit_release_ral / hatch_release,
    ral_exp_rate  = 1 / ral_mark_rate
  ) %>%
  rename(
    rel_group = hatchery,
    rel_year = migr_year
  )

```

```{r krs-check, eval = FALSE}
krs_check = sfsr_comp_obs_df_yr %>%
  filter(
    node %in% c("KRS_D", "KRS_U"),
    rel_site == "LGRLDR"
  ) %>%
  mutate(age = if_else(length < jack_cutoff_mm, "jack", "adult")) %>%
  group_by(age) %>%
  summarize(n = n_distinct(tag_code), .groups = "drop")
  
```

```{r prep-data}
# summarize tags by release group, release year, ocean age, sort-by-code designation, expansion rate, and array
tag_df = sfsr_comp_obs_df_yr %>%
  select(
    tag_code,
    spawn_yr,
    node,
    event_type_name,
    min_det,
    mark_site,
    mark_date,
    mark_rear_type_name,
    rel_site,
    rel_date,
    length,
    flags
  ) %>%
  mutate(rel_year = year(rel_date)) %>%
  filter(node %in% nodes_all) %>%
  # for any tag detected at a node multiple times, just keep first detection
  group_by(tag_code, node) %>%
  slice_min(min_det, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  # join tag_specific lgr tag expansion rates
  left_join(lgr_tag_exp, by = c("tag_code" = "tag")) %>%
  mutate(
    ocean_age = case_when(
      rel_site == "LGRLDR" & length <  jack_cutoff_mm        ~ 1,
      rel_site == "LGRLDR" & length <  adult_2salt_cutoff_mm ~ 2,
      rel_site == "LGRLDR" & length >= adult_2salt_cutoff_mm ~ 3,
      TRUE ~ ocean_age
    ),
    rel_group = assign_rel_group(
      rel_site            = rel_site,
      flags               = flags,
      mark_rear_type_name = mark_rear_type_name
    )
  ) %>%
  # only interested in fish tagged at LGR or hatchery releases from MCCA sites in SF Salmon
  filter(
    mark_site %in% c("LGRLDR", "MCCA"),
    !is.na(rel_group)
  ) %>%
  group_by(
    rel_group,
    rel_year,
    ocean_age,
    sby_c,
    exp_rate,
    node
  ) %>%
  summarize(
    n_tags = n_distinct(tag_code),
    .groups = "drop"
  ) %>%
  arrange(
    rel_group,
    rel_year,
    node
  ) %>%
  # join bonneville group mark rates as backup when tag-specific expansions unavailable
  left_join(
    bon_exp_rates %>%
      select(rel_group, rel_year, ral_exp_rate),
    by = c("rel_group", "rel_year")
  ) %>%
  # apply lgr expansion rate to LGRLDR rel_group
  mutate(
    exp_rate = case_when(
      rel_group == "LGRLDR" ~ 1 / lgr_trap_rate,
      is.na(exp_rate)       ~ ral_exp_rate,
      TRUE                  ~ exp_rate
    ),
    n_tags_exp = round(n_tags * exp_rate)
  ) %>%
  select(-ral_exp_rate)

```

```{r tag-exp-mark-rate}
tag_exp = tag_df %>%
  group_by(
    rel_group,
    rel_year,
    ocean_age,
    node
  ) %>%
  summarise(
    n_tags     = sum(n_tags, na.rm = TRUE),
    n_tags_exp = sum(n_tags_exp, na.rm = TRUE),
    .groups = "drop"
  )

```

```{r det-probs, results = "hide"}
load(here("data/configuration_files/site_config_LGR_20260116.rda")) ; rm(flowlines, crb_sites_sf, sr_site_pops)

node_paths = addParentChildNodes(parent_child, configuration) %>%
  buildNodeOrder()

# calculate node detection efficiency, all spawn years
node_p = estNodeEff(
  capHist_proc = sfsr_comp_obs_df_yr,
  node_order   = node_paths
) %>%
  filter(node %in% nodes_all)

```

```{r tag-expansion}
exp_df = tag_exp %>%
  filter(str_detect(rel_group, "McCall") | str_detect(rel_group, "LGR")) %>%
  left_join(
    node_p %>%
      select(node, eff_est, eff_se),
    by = "node"
  ) %>%
  filter(!(node %in% nodes_upstream)) %>%
  mutate(
    age = case_when(
      ocean_age == 1            ~ "Jacks",
      ocean_age %in% c(2, 3, 4) ~ "Adults",
      TRUE                      ~ NA_character_
    ),
    est_tags = round(n_tags     / eff_est),
    tot_est  = round(n_tags_exp / eff_est)
  ) %>%
  group_by(
    rel_group,
    age,
    node,
    eff_est
  ) %>%
  summarize(
    n_tags     = sum(n_tags, na.rm = TRUE),
    est_tags   = sum(est_tags, na.rm = TRUE),
    n_tags_exp = sum(n_tags_exp, na.rm = TRUE),
    tot_est    = sum(tot_est, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  select(
    rel_group,
    age,
    node,
    n_tags,
    eff_est,
    est_tags,
    n_tags_exp,
    tot_est
  ) %>%
  arrange(
    rel_group,
    desc(node),
    age
  )

```

**Data Last Retrieved:** `r format(update_dttm, "%B %d, %Y %I:%M %p %Z")`

To estimate the number of adults past each array (Table \@ref(tab:results-tbl)), we summarized expanded counts by release group and age (i.e., adults versus jacks) and further accounted for array detection efficiencies. Array detection efficiencies were estimated as the percent of tags at upstream array(s) that were detected at the given array.

PIT tag observations were queried from DART at the above date and time and summarized by release group, release year, and sort-by-code-designation (i.e., run-at-large [RAL] versus return-to-river [RTR]) to determine the number of tags observed at the South Fork Guard Station (SFS) and Krassel (KRS) arrays on the South Fork Salmon River (Table \@ref(tab:tag-tbl)). The number of tags observed were then expanded based on expansion rate, calculated as the inverse of mark rates. For in-season escapement estimates, we focus on natural-origin adults at Lower Granite Dam and juvenile hatchery releases to the South Fork Salmon River. For hatchery releases, expansion rates were derived from the current-year **PIT Data** tab in `r basename(lgr_mark_filepath)`. Natural-origin expansion rates were assumed to be `r round(1 / lgr_trap_rate, 2)` based on the set trapping and tagging rate at LGR (`r round(lgr_trap_rate * 100, 1)`%).

```{r results-tbl, tab.cap = "Estimated number of adults (jacks excluded) past each array, by release group, after expanding by mark rates and array detection efficiencies."}
results_tbl = exp_df %>%
  filter(age == "Adults") %>%
  filter(node %in% nodes_report) %>%
  mutate(node = factor(node, levels = nodes_report)) %>%
  select(-n_tags_exp) %>%
  mutate(eff_est = round(eff_est * 100)) %>%
  rename(
    `Release Group`     = rel_group,
    Age                 = age,
    Array               = node,
    `Tags Observed`     = n_tags,
    `Array Det Eff (%)` = eff_est,
    `Est Tags`          = est_tags,
    `Est Adults`        = tot_est
  ) %>%
  arrange(`Release Group`, Array, Age) %>%
  kable(
    booktabs = TRUE,
    align = rep("c", ncol(.)),
    escape = FALSE,
    label = NA
  ) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "condensed")
  )

results_tbl

```

```{r run-timing-calcs}
sfsr_fish = sfsr_comp_obs_df %>%
  filter(
    mark_site %in% c("LGRLDR", "MCCA"),
    rel_site  %in% c("LGRLDR", "KNOXB")
  ) %>%
  mutate(
    node = recode(node, KRS = "KRS_D"),
    julian = mdy(paste0(format(min_det, format = "%m-%d"), "-2020")),
    rel_group = assign_rel_group(
      rel_site = rel_site,
      flags = flags,
      mark_rear_type_name = mark_rear_type_name
    )
  )

obs = sfsr_fish %>%
  filter(
    !is.na(rel_group),
    node %in% c("SFG", "KRS_D")
  ) %>%
  group_by(rel_group, node, tag_code) %>%
  slice_min(min_det, n = 1, with_ties = FALSE) %>%
  ungroup()

obs_sum = obs %>%
  group_by(
    rel_group, 
    node, 
    spawn_yr, 
    julian
  ) %>%
  summarise(n = n(), .groups = "drop") %>%
  complete(rel_group, node, spawn_yr, julian, fill = list(n = 0)) %>%
  group_by(rel_group, node, spawn_yr) %>%
  mutate(
    csum = cumsum(n),
    cdf  = csum / sum(n)
  ) %>%
  ungroup()

today = mdy(paste0(format(Sys.Date(), format = "%m-%d"), "-2020"))

avg_run = obs_sum %>%
  filter(spawn_yr != current_year) %>%
  group_by(
    rel_group, 
    node, 
    julian) %>%
  summarise(cdf = mean(cdf, na.rm = TRUE), .groups = "drop")

run_timing_today = obs_sum %>%
  filter(julian == today) %>%
  group_by(
    rel_group, 
    node) %>%
  summarise(cdf = mean(cdf, na.rm = TRUE), .groups = "drop")

```

```{r run-timing, fig.align = "center", fig.cap = paste0("Cumulative run timing of hatchery integrated, hatchery segregated, and unclipped returns of LGRLDR-released tags across South Fork Salmon River PIT-tag arrays. Historic returns from 2010 - ", current_year - 1, " (grey) and their average run timing (blue) provide context for the current ", current_year, " return (green). The vertical dashed red line indicates the last data download.")}
run_time_p = ggplot() +
  geom_line(
    data = obs_sum %>% filter(spawn_yr != current_year),
    aes(x = julian, y = cdf, group = spawn_yr),
    colour = "grey"
  ) +
  geom_line(
    data = avg_run,
    aes(x = julian, y = cdf),
    colour = "navy",
    linewidth = 1
  ) +
  geom_line(
    data = obs_sum %>% filter(spawn_yr == current_year),
    aes(x = julian, y = cdf),
    colour = "limegreen"
  ) +
  geom_vline(
    xintercept = today,
    colour = "firebrick",
    linetype = "dashed"
  ) +
  scale_x_date(
    date_breaks = "3 weeks",
    date_labels = "%b-%d"
  ) +
  facet_grid(rel_group ~ fct_rev(node)) +
  theme_bw() +
  labs(
    x = NULL,
    y = "CDF"
  )

run_time_p

```

```{r run-timing-year, fig.align = "center", fig.height = 9, fig.cap = paste0("Cumulative run timing of PIT-tagged spring-summer Chinook salmon detected at South Fork Salmon River arrays by spawn year. Curves are shown separately for natural-origin (LGRLDR), hatchery integrated, and hatchery segregated. The vertical dashed red line indicates the last data download.")}
run_time_yr_p = ggplot() +
  stat_ecdf(
    data = obs,
    aes(x = julian, colour = rel_group),
    linewidth = 1
  ) +
  geom_vline(
    xintercept = today,
    colour = "firebrick",
    linetype = "dashed"
  ) +
  scale_x_date(
    date_breaks = "4 weeks",
    date_labels = "%b-%d"
  ) +
  facet_wrap(~ spawn_yr, ncol = 2) +
  theme_bw() +
  theme(legend.position = "top") +
  labs(
    x = NULL,
    y = "CDF",
    colour = ""
  )

run_time_yr_p
```

```{r site-arrival-p, eval = T, fig.align = "center", fig.height = 9, fig.cap = "Daily counts of PIT-tagged adult Chinook salmon first detected at LGR and South Fork Salmon River IPTDS sites during the current spawn year. Bars show the number of tagged fish first detected each day, stacked by release group. Faint dashed lines and shaded ribbons represent the mean and interquartile range (25th-75th percentile), respectively, for each release group in previous years since 2020. The vertical dashed red line indicates the current date."}

site_levels = c("LGR", "SFG", "KRS", "ESS", "ZEN")

start_date = as.Date("2020-05-10")
end_date   = as.Date("2020-08-01")
today_2020 = as.Date(sprintf("2020-%s", format(Sys.Date(), "%m-%d")))

daily_dat = sfsr_fish %>%
  filter(
    spawn_yr >= 2020,
    !is.na(rel_group)
  ) %>%
  mutate(
    julian = as.Date(julian),
    site = case_when(
      str_detect(node, "^LGR") ~ "LGR",
      str_detect(node, "^SFG") ~ "SFG",
      str_detect(node, "^KRS") ~ "KRS",
      str_detect(node, "^ESS") ~ "ESS",
      str_detect(node, "^ZEN") ~ "ZEN",
      TRUE ~ NA_character_
    ),
    site = factor(site, levels = site_levels)
  ) %>%
  filter(!is.na(site)) %>%
  group_by(spawn_yr, tag_code, rel_group, site) %>%
  summarise(
    first_det = min(julian, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(
    first_det >= start_date,
    first_det <= end_date
  ) %>%
  count(spawn_yr, rel_group, site, first_det, name = "n_tags") %>%
  complete(
    spawn_yr,
    site,
    rel_group,
    first_det = seq.Date(start_date, end_date, by = "day"),
    fill = list(n_tags = 0)
  )

hist_dat = daily_dat %>%
  filter(spawn_yr != current_year) %>%
  group_by(rel_group, site, first_det) %>%
  summarise(
    mean_tags = mean(n_tags, na.rm = TRUE),
    q25 = quantile(n_tags, 0.25, na.rm = TRUE),
    q75 = quantile(n_tags, 0.75, na.rm = TRUE),
    lo = pmin(q25, mean_tags),
    hi = pmax(q75, mean_tags),
    .groups = "drop"
  )

current_dat = daily_dat %>%
  filter(spawn_yr == current_year)

site_arrival_p = ggplot() +
  geom_ribbon(
    data = hist_dat,
    aes(
      x = first_det,
      ymin = lo,
      ymax = hi,
      group = interaction(site, rel_group)
    ),
    fill = "grey70",
    alpha = 0.12,
    color = NA
  ) +
  geom_line(
    data = hist_dat,
    aes(
      x = first_det,
      y = mean_tags,
      colour = rel_group,
      group = interaction(site, rel_group)
    ),
    linewidth = 0.6,
    alpha = 0.4,
    linetype = "22",
    lineend = "round"
  ) +
  geom_col(
    data = current_dat,
    aes(
      x = first_det,
      y = n_tags,
      fill = rel_group
    ),
    width = 0.95,
    colour = "white",
    linewidth = 0.15
  ) +
  geom_vline(
    xintercept = today_2020,
    colour = "firebrick",
    linetype = "dashed",
    linewidth = 0.6
  ) +
  facet_grid(site ~ ., scales = "free_y") +
  scale_x_date(
    limits = c(start_date, end_date),
    date_breaks = "1 weeks",
    date_labels = "%b %d",
    expand = expansion(mult = c(0.005, 0.01))
  ) +
  scale_fill_viridis_d(
    option = "D",
    end = 0.9,
    name = "Release group"
  ) +
  scale_color_viridis_d(
    option = "D",
    end = 0.9,
    guide = "none"
  ) +
  labs(
    x = NULL,
    y = "Observed Tags"
  ) +
  theme_bw(base_size = 12) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_line(colour = "grey92"),
    strip.background = element_rect(fill = "grey90", colour = "grey60"),
    strip.text.y = element_text(face = "bold", angle = 0),
    legend.position = "bottom",
    #legend.title = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.spacing = unit(0.15, "lines")
  )

site_arrival_p

```

```{r tag-tbl, tab.cap = "PIT tags detected at South Fork Salmon River arrays by release group, release year, sort-by-code designation, and ocean age. Expanded estimates of adult escapement at each array were calculated from observed tags and expansion rates but were not adjusted for array detection efficiency."}
tag_tbl = tag_df %>%
  filter(node %in% nodes_report) %>%
  mutate(node = factor(node, levels = nodes_report)) %>%
  select(
    rel_group,
    rel_year,
    sby_c,
    node,
    ocean_age,
    n_tags,
    exp_rate,
    n_tags_exp
  ) %>%
  arrange(
    rel_group,
    node,
    rel_year
  ) %>%
  rename(
    `Release Group`      = rel_group,
    `Release Year`       = rel_year,
    `Sort-By-Code`       = sby_c,
    Array                = node,
    `Ocean Age`          = ocean_age,
    `Tags Observed`      = n_tags,
    `Expansion Rate`     = exp_rate,
    `Mark Rate Expanded` = n_tags_exp
  ) %>%
  kable(
    booktabs = TRUE,
    align = "ccccccc",
    label = NA
  ) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "condensed")
  )

tag_tbl

```

***END DOCUMENT***
<!-- End document -->