---
title: "In-season Escapement Estimates to SFG and KRS PIT-tag Arrays in the South Fork Salmon River"
author:
  - name: "Mike Ackerman"
    email: mikea@nezperce.org
    affiliation: "Research Scientist"
    correspondence: yes

  - name: "Ryan Kinzer"
    affiliation: "Fisheries Data Analysis Coordinator"

  - name: "Jason Vogel"
    affiliation: "Research Division Director"

date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    theme: cerulean
---

<!-- the following inserts the NPT logo into header -->
```{=html}
<script>
   $(document).ready(function() {
     $head = $('#header');
     $head.prepend('<img src=\"DFRM.png\" style=\"float: right;width: 150px;\"/>')
   });
</script>

<style>
p.caption {
  font-size: 0.8em;
}
</style>
```

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE
)
```

```{r load-libraries}
# for formatting
library(kableExtra)

# load necessary libraries
library(tidyverse)
library(lubridate)
library(readxl)
library(here)
library(janitor)
library(PITcleanr)
```

```{r load-obs-data}
# load latest compiled observation dataset based on dttm
load(here("data/sfsr_chnk_compiled_obs_latest.rda"))
current_year = year(update_dttm)
# current_year = 2025

# filter to current_year
sfsr_comp_obs_df_yr = sfsr_comp_obs_df %>%
  filter(spawn_yr == current_year)
```

```{r set-parameters}
# lgr trap and subsample rates
lgr_trap_rate         = 0.186       # assumed lgr trapping/tagging rate
lgr_hor_chnk_sub_rate = 0.33333333  # assumed lgr hatchery chinook subsample rate

# chnk size cutoffs
jack_cutoff_mm        = 580
adult_2salt_cutoff_mm = 815

# paths to mark/expansion rates
lgr_mark_filepath     = here("data/mark_rates/2026 PIT_Tag_Analysis_LowerGranite.xlsx") # path to lgr tag-specific expansion data
bon_exp_filepath      = here("data/mark_rates/2026 PIT_Tag_Analysis_Bonneville.xlsx")   # path to past bon expansion rates (backup option)
# lgr_mark_filepath = here("data/mark_rates/deprecated/2025 PIT_Tag_Analysis_LowerGranite.xlsx") 
# bon_exp_filepath  = here("data/mark_rates/deprecated/2025 PIT_Tag_Analysis_Bonneville.xlsx")

# array sets
nodes_all      = c("SFG", "KRS", "KRS_D", "KRS_U", "STR", "ESS_D", "ESS_U", "ZEN_D", "ZEN_U")
nodes_report   = c("SFG", "KRS_D", "ESS_D", "ZEN_D")
nodes_upstream = c("KRS_U", "ESS_U", "ZEN_U")

```

```{r assign-rg-helper}
assign_rel_group = function(rel_site,
                            flags,
                            mark_rear_type_name) {
  case_when(
    rel_site == "KNOXB"  & str_detect(flags, "AI")    ~ "McCall - Integrated",
    rel_site == "KNOXB"  & str_detect(flags, "AD")    ~ "McCall - Segregated",
    rel_site == "LGRLDR" & mark_rear_type_name == "W" ~ "LGRLDR",
    rel_site == "LGRLDR" & str_detect(flags, "CW")    ~ "LGRLDR",
    TRUE ~ NA_character_
  )
}
```

```{r load-exp-data}
# load lgr tag expansions, the preferred option, for current year
lgr_tag_exp = read_excel(path = lgr_mark_filepath,
                         sheet = "PIT Data") %>%
  clean_names() %>%
  filter(!is.na(tag)) %>%
  select(
    tag,
    brood_year,
    release_site_code,
    ocean_age,
    sby_c,
    exp_rate = expansion
  )

# load historic, summarized mark rates from bonneville, the backup option
bon_exp_rates = read_excel(path = bon_exp_filepath,
                           sheet = "Historic Juv Rel Numbers") %>%
  clean_names() %>%
  filter(str_detect(hatchery, "McCall")) %>%
  mutate(
    hatchery = case_when(
      hatchery == "McCall (Int)" ~ "McCall - Integrated",
      hatchery == "McCall (Seg)" ~ "McCall - Segregated",
      TRUE ~ hatchery
    )
  ) %>%
  mutate(
    ral_mark_rate = pit_release_ral / hatch_release,
    ral_exp_rate  = 1 / ral_mark_rate
  ) %>%
  rename(
    rel_group = hatchery,
    rel_year = migr_year
  )

```

```{r krs-check}
krs_check = sfsr_comp_obs_df_yr %>%
  filter(
    node %in% c("KRS_D", "KRS_U"),
    rel_site == "LGRLDR"
  ) %>%
  mutate(age = if_else(length < jack_cutoff_mm, "jack", "adult")) %>%
  group_by(age) %>%
  summarize(n = n_distinct(tag_code), .groups = "drop")
  
```

```{r prep-data}
# summarize tags by release group, release year, ocean age, sort-by-code designation, expansion rate, and array
tag_df = sfsr_comp_obs_df_yr %>%
  select(
    tag_code,
    spawn_yr,
    node,
    event_type_name,
    min_det,
    mark_site,
    mark_date,
    mark_rear_type_name,
    rel_site,
    rel_date,
    length,
    flags
  ) %>%
  mutate(rel_year = year(rel_date)) %>%
  filter(node %in% nodes_all) %>%
  # for any tag detected at a node multiple times, just keep first detection
  group_by(tag_code, node) %>%
  slice_min(min_det, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  # join tag_specific lgr tag expansion rates
  left_join(lgr_tag_exp, by = c("tag_code" = "tag")) %>%
  mutate(
    ocean_age = case_when(
      rel_site == "LGRLDR" & length <  jack_cutoff_mm        ~ 1,
      rel_site == "LGRLDR" & length <  adult_2salt_cutoff_mm ~ 2,
      rel_site == "LGRLDR" & length >= adult_2salt_cutoff_mm ~ 3,
      TRUE ~ ocean_age
    ),
    rel_group = assign_rel_group(
      rel_site            = rel_site,
      flags               = flags,
      mark_rear_type_name = mark_rear_type_name
    )
  ) %>%
  # only interested in fish tagged at LGR or hatchery releases from MCCA sites in SF Salmon
  filter(
    mark_site %in% c("LGRLDR", "MCCA"),
    !is.na(rel_group)
  ) %>%
  group_by(
    rel_group,
    rel_year,
    ocean_age,
    sby_c,
    exp_rate,
    node
  ) %>%
  summarize(
    n_tags = n_distinct(tag_code),
    .groups = "drop"
  ) %>%
  arrange(
    rel_group,
    rel_year,
    node
  ) %>%
  # join bonneville group mark rates as backup when tag-specific expansions unavailable
  left_join(
    bon_exp_rates %>%
      select(rel_group, rel_year, ral_exp_rate),
    by = c("rel_group", "rel_year")
  ) %>%
  # apply
  mutate(
    exp_rate = case_when(
      rel_group == "LGRLDR" ~ 1 / lgr_trap_rate,
      is.na(exp_rate)       ~ ral_exp_rate,
      TRUE                  ~ exp_rate
    ),
    n_tags_exp = round(n_tags * exp_rate)
  ) %>%
  select(-ral_exp_rate)

```

```{r tag-exp-mark-rate}
tag_exp = tag_df %>%
  group_by(
    rel_group,
    rel_year,
    ocean_age,
    node
  ) %>%
  summarise(
    n_tags     = sum(n_tags, na.rm = TRUE),
    n_tags_exp = sum(n_tags_exp, na.rm = TRUE),
    .groups = "drop"
  )

```

```{r det-probs, results = "hide"}
load(here("data/configuration_files/site_config_LGR_20260116.rda")) ; rm(flowlines, crb_sites_sf, sr_site_pops)

node_paths = addParentChildNodes(parent_child, configuration) %>%
  buildNodeOrder()

# calculate node detection efficiency, all spawn years
node_p = estNodeEff(
  capHist_proc = sfsr_comp_obs_df_yr,
  node_order   = node_paths
) %>%
  filter(node %in% nodes_all)

```

```{r tag-expansion}
exp_df = tag_exp %>%
  filter(str_detect(rel_group, "McCall") | str_detect(rel_group, "LGR")) %>%
  left_join(
    node_p %>%
      select(node, eff_est, eff_se),
    by = "node"
  ) %>%
  filter(!(node %in% nodes_upstream)) %>%
  mutate(
    age = case_when(
      ocean_age == 1         ~ "Jacks",
      ocean_age %in% c(2, 3) ~ "Adults",
      TRUE                   ~ NA_character_
    ),
    est_tags = round(n_tags / eff_est),
    tot_est  = round(n_tags_exp / eff_est)
  ) %>%
  group_by(
    rel_group,
    age,
    node,
    eff_est,
    eff_se
  ) %>%
  summarize(
    n_tags     = sum(n_tags, na.rm = TRUE),
    est_tags   = sum(est_tags, na.rm = TRUE),
    n_tags_exp = sum(n_tags_exp, na.rm = TRUE),
    tot_est    = sum(tot_est, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  select(
    rel_group,
    age,
    node,
    n_tags,
    eff_est,
    eff_se,
    est_tags,
    n_tags_exp,
    tot_est
  ) %>%
  arrange(
    rel_group,
    desc(node),
    age
  )

```

**Data Last Retrieved:** `r format(update_dttm, "%B %d, %Y %I:%M %p %Z")`

To estimate the number of adults past each array (Table \@ref(tab:results-tbl)), we summarized expanded counts by release group and age (i.e., adults versus jacks) and further accounted for array detection efficiencies. Array detection efficiencies were estimated as the percent of tags at upstream array(s) that were detected at the given array.

PIT tag observations were queries from DART at the above date and time and summarized by release group, release year, and sort-by-code-designation (i.e., run-at-large [RAL] versus return-to-river [RTR]) to determine the number of tags observed at the South Fork Guard Station (SFS) and Krassel (KRS) arrays on the South fork Salmon River (Table \@ref(tab:tag-tbl)). The number of tags observed were then expanded based on expansion rate, calculated as the inverse of mark rates. For in-season escapement estimates, we focus on natural-origin adults at Lower Granite Dam and juvenile hatchery releases to the South Fork Salmon River. For hatchery releaseds, expansion rates were derived from the current-year **PIT Data** tab in `r basename(lgr_mark_filepath)`. Natural-origin expansion rates were assumed to be `r round(1 / lgr_trap_rate, 2)` based on the set trapping and tagging rate at LGR (`r round(lgr_trap_rate * 100, 1)`%).

```{r results-tbl, tab.cap = "Estimated number of adults past each array by release group and age after expanding by mark rates and array detection efficiencies."}
results_tbl = exp_df %>%
  filter(age == "Adults") %>%
  filter(node %in% nodes_report) %>%
  mutate(node = factor(node, levels = nodes_report)) %>%
  select(-eff_se, -n_tags_exp) %>%
  mutate(eff_est = round(eff_est * 100)) %>%
  rename(
    `Release Group` = rel_group,
    Age = age,
    Array = node,
    `Tags Observed` = n_tags,
    `Array Det Eff` = eff_est,
    `Estimated Tags` = est_tags,
    `Est. Adults` = tot_est
  ) %>%
  arrange(`Release Group`, Array, Age) %>%
  kable(
    booktabs = TRUE,
    align = "ccccccc",
    label = NA
  ) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "condensed")
  )

results_tbl

```

```{r run-timing-calcs}
sfsr_fish = sfsr_comp_obs_df %>%
  filter(
    mark_site %in% c("LGRLDR", "MCCA"),
    rel_site %in% c("LGRLDR", "KNOXB")
  ) %>%
  mutate(
    node = recode(node, KRS = "KRS_D"),
    julian = mdy(paste0(format(min_det, format = "%m-%d"), "-2020")),
    rel_group = assign_rel_group(
      rel_site = rel_site,
      flags = flags,
      mark_rear_type_name = mark_rear_type_name
    )
  )

obs = sfsr_fish %>%
  filter(
    !is.na(rel_group),
    node %in% c("SFG", "KRS_D")
  ) %>%
  group_by(rel_group, node, tag_code) %>%
  slice_min(min_det, n = 1, with_ties = FALSE) %>%
  ungroup()

obs_sum = obs %>%
  group_by(
    rel_group, 
    node, 
    spawn_yr, 
    julian
  ) %>%
  summarise(n = n(), .groups = "drop") %>%
  complete(rel_group, node, spawn_yr, julian, fill = list(n = 0)) %>%
  group_by(rel_group, node, spawn_yr) %>%
  mutate(
    csum = cumsum(n),
    cdf  = csum / sum(n)
  ) %>%
  ungroup()

today = mdy(paste0(format(Sys.Date(), format = "%m-%d"), "-2020"))

avg_run = obs_sum %>%
  filter(spawn_yr != current_year) %>%
  group_by(
    rel_group, 
    node, 
    julian) %>%
  summarise(cdf = mean(cdf, na.rm = TRUE), .groups = "drop")

run_timing_today = obs_sum %>%
  filter(julian == today) %>%
  group_by(
    rel_group, 
    node) %>%
  summarise(cdf = mean(cdf, na.rm = TRUE), .groups = "drop")

```

```{r run-timing, fig.align = "center", fig.cap = paste0("Run timing of hatchery integrated, hatchery segregated, and unclipped returns of LGRLDR-released tags across South Fork Salmon River PIT-tag arrays. Grey lines indicate individual spawn year returns before ", current_year, ", the dark blue line is the average cumulative proportion of returns before ", current_year, ", and the green line represents the current ", current_year, " return.")}
run_time_p = ggplot() +
  geom_line(
    data = obs_sum %>% filter(spawn_yr != current_year),
    aes(x = julian, y = cdf, group = spawn_yr),
    colour = "grey"
  ) +
  geom_line(
    data = avg_run,
    aes(x = julian, y = cdf),
    colour = "navy",
    linewidth = 1
  ) +
  geom_line(
    data = obs_sum %>% filter(spawn_yr == current_year),
    aes(x = julian, y = cdf),
    colour = "limegreen"
  ) +
  geom_vline(
    xintercept = today,
    colour = "firebrick",
    linetype = "dashed"
  ) +
  scale_x_date(
    date_breaks = "3 weeks",
    date_labels = "%b-%d"
  ) +
  facet_grid(rel_group ~ fct_rev(node)) +
  theme_bw() +
  labs(
    x = "Date",
    y = "CDF"
  )

run_time_p

```

```{r run-timing-year, fig.align = "center", fig.height = 9, fig.cap = "Run timing of hatchery integrated, hatchery segregated, and unclipped returns of LGRLDR-released tags across South Fork Salmon River PIT-tag arrays. The vertical red line indicates the last data download."}
run_time_yr_p = ggplot() +
  stat_ecdf(
    data = obs,
    aes(x = julian, colour = rel_group),
    linewidth = 1
  ) +
  geom_vline(
    xintercept = today,
    colour = "firebrick",
    linetype = "dashed"
  ) +
  scale_x_date(
    date_breaks = "4 weeks",
    date_labels = "%b-%d"
  ) +
  facet_wrap(~ spawn_yr, ncol = 2) +
  theme_bw() +
  theme(legend.position = "top") +
  labs(
    x = "Date",
    y = "CDF",
    colour = ""
  )

run_time_yr_p
```

```{r tag-tbl, tab.cap = "PIT tags detected at South Fork Salmon River arrays by release group, release year, and sort-by-code designation including expansion rates and an estimate of the number of adults at each array, not accounting for array detection probability."}
tag_tbl = tag_df %>%
  filter(node %in% nodes_report) %>%
  mutate(node = factor(node, levels = nodes_report)) %>%
  select(
    rel_group,
    rel_year,
    sby_c,
    node,
    n_tags,
    exp_rate,
    n_tags_exp
  ) %>%
  arrange(
    rel_group,
    node,
    rel_year
  ) %>%
  rename(
    `Release Group` = rel_group,
    `Release Year` = rel_year,
    `Sort-By-Code` = sby_c,
    Array = node,
    `Tags Observed` = n_tags,
    `Expansion Rate` = exp_rate,
    `Mark Rate Expanded` = n_tags_exp
  ) %>%
  kable(
    booktabs = TRUE,
    align = "ccccccc",
    label = NA
  ) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "condensed")
  )

tag_tbl

```

<!-- End document -->