---
title: "In-season Escapement Estimates to the SFG and KRS PIT-tag Arrays in the South Fork Salmon River"
author:
  - name: "Michael W. Ackerman"
    email: mikea@nezperce.org
    affiliation: "Research Scientist"
    correspondence: yes

  - name: "Ryan N. Kinzer"
    affiliation: "Fisheries Data Analysis Coordinator"

  - name: "Jason Vogel"
    affiliation: "Research Division Director"

date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    theme: cerulean
---

<!-- the following inserts the NPT logo into header -->
```{=html}
<script>
   $(document).ready(function() {
     $head = $('#header');
     $head.prepend('<img src=\"DFRM.png\" style=\"float: right;width: 150px;\"/>')
   });
</script>

<style>
p.caption {
  font-size: 0.8em;
}
</style>
```

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE
  # comment = "#>",
  # fig.path = "../figures/",
  # dev = c("pdf", "png", "tiff"),
  # dpi = 400
)

```

```{r load-libraries}
# for formatting
library(kableExtra)

# load necessary libraries
library(tidyverse)
library(here)
library(readxl)
library(janitor)

```

```{r set-parameters}
yr = 2025                                                 # current year
lgr_trap_rate = 0.20                                      # set and assumed lgr trapping and tagging rate
lgr_hor_subsample_rate = 0.25
lgr_mark_filepath = here("data/mark_rates/2025 PIT_Tag_Analysis_LowerGranite.xlsx")  # path to lgr tag-specific expansion data
bon_exp_filepath  = here("data/mark_rates/2025 PIT_Tag_Analysis_Bonneville.xlsx")    # path to past bon expansion rates (backup option)

```

```{r load-data}
# get the date-time from the most recent dart observations
dt_tm = list.files(path = here("data/observations/"),
                   recursive = F,
                   full.names = F) %>%
  sort() %>%
  .[length(.)] %>%
  sub("^[^_]*_[^_]*_([^\\.]+)\\.rda$", "\\1", .)

# load dart observation data
load(paste0(here("data/observations/sfsr_obs_"), dt_tm, ".rda"))

# filter to year of interest
sfsr_obs_yr = sfsr_obs %>%
  filter(spawn_year == yr)

# load lgr tag expansion data, the preferred option, for current year
lgr_tag_exp = read_excel(path = lgr_mark_filepath,
                         sheet = "PIT Data") %>%
  clean_names() %>%
  filter(!is.na(tag)) %>%
  select(tag,
         release_site_code,
         brood_year,
         ocean_age,
         sby_c,
         exp_rate = expansion)

# load historic generic mark rates from bon, the backup option
bon_exp_rates = read_excel(path = bon_exp_filepath,
                        sheet = "Historic Juv Rel Numbers") %>%
  clean_names() %>%
  filter(str_detect(hatchery, "McCall")) %>%
  mutate(hatchery = case_when(
    hatchery == "McCall (Int)" ~ "McCall - Integrated",
    hatchery == "McCall (Seg)"  ~ "McCall - Segregated",
    TRUE ~ hatchery
  )) %>%
  mutate(ral_mark_rate = pit_release_ral / hatch_release,
         ral_exp_rate = 1 / ral_mark_rate) %>%
  rename(rel_group = hatchery,
         rel_year = migr_year)

```

```{r krs-only}
jack_cutoff <- 580
krs <- sfsr_obs_yr %>%
  filter(node %in% c('KRS_D', 'KRS_U')) %>%
  filter(rel_site == 'LGRLDR') %>%
  mutate(age = ifelse(length < jack_cutoff, 'jack', 'adult')) %>%
  group_by(age) %>%
  summarize(n = n_distinct(tag_code))
```


```{r prep-data}
# summarize tags by release group, release year, and site code (node)
tag_df = sfsr_obs_yr %>%
  select(spawn_year,
         tag_code,
         node,
         event_type_name,
         min_det,
         mark_site,
         mark_date,
         mark_rear_type_name,
         rel_site,
         rel_date,
         length,
         flags) %>%
  mutate(rel_year = year(rel_date)) %>%
  # for any tag that was observed at a node multiple times, just keep the first detection
  group_by(tag_code, node) %>%
  filter(min_det == min(min_det)) %>%
  ungroup() %>%
  # keep just detections within mainsten SF Salmon River
  filter(node %in% c("SFG", "KRS_D", "KRS_U", "STR", "ESS_D", "ESS_U", "ZEN_D", "ZEN_U")) %>%
  # join tag-specific lgr tag detection rates
  left_join(lgr_tag_exp, by = c("tag_code" = "tag")) %>%
  mutate(ocean_age = case_when(
    rel_site == 'LGRLDR' & length < jack_cutoff ~ 1,
    rel_site == 'LGRLDR' & length < 815 ~ 2,
    rel_site == 'LGRLDR' & length >= 815 ~ 3,
    TRUE ~ ocean_age)) %>%
  group_by(mark_site,
           mark_rear_type_name,
           rel_site,
           rel_year,
           ocean_age,
           sby_c,
           exp_rate,
           flags,
           node) %>%
  summarise(n = n_distinct(tag_code),
            .groups = "drop") %>%
  # only interested in fish tagged at LGR or hatchery releases from MCCA sites in SF Salmon
  filter(mark_site %in% c("LGRLDR", "MCCA")) %>%
  mutate(rel_group = case_when(
    rel_site == "KNOXB" & str_detect(flags, "AI")     ~ "McCall - Integrated",
    rel_site == "KNOXB" & str_detect(flags, "AD")     ~ "McCall - Segregated",
    rel_site == "LGRLDR" & mark_rear_type_name == "W" ~ "LGRLDR", # - NOR",
    rel_site == "LGRLDR" & str_detect(flags, "CW")    ~ "LGRLDR", # - HOR (AD-intact, CWT)",
    #rel_site == "KNOXB"  & mark_rear_type_name == "W" ~ "KNOXB - NOR",
    #rel_site == "SFSRKT" & mark_rear_type_name == "W" ~ "SFSRKT - NOR",
    TRUE ~ NA
  )) %>%
  group_by(rel_group,
           rel_year,
           ocean_age,
           sby_c,
           exp_rate,
           node) %>%
  summarize(n_tags = sum(n),
            .groups = "drop") %>%
  arrange(rel_group,
          rel_year,
          node) %>%
  # join bon group mark rates (back-up)
  left_join(bon_exp_rates %>%
              select(rel_group,
                     rel_year,
                     ral_exp_rate),
            by = c("rel_group", "rel_year")) %>%
  # lgr expansion rate
  mutate(exp_rate = case_when(
    str_detect(rel_group, "LGRLDR") ~ 1 / lgr_trap_rate,                            # the LGR sample rate
    #str_detect(rel_group, fixed("LGR - HOR (AD-intact, CWT)")) ~ 1 / lgr_trap_rate, # * lgr_hor_subsample_rate),
    is.na(exp_rate) ~ ral_exp_rate,                                                    # if not lgr expansion rate, use bon rate
    TRUE ~ exp_rate
  )) %>%
  select(-ral_exp_rate) %>%
  mutate(n_tags_exp = round(n_tags * exp_rate))

```

```{r tag-exp-mark-rate}
tag_exp = tag_df %>%
  # mutate(exp_rate = case_when(
  #   str_detect(rel_group, "LGR - NOR") ~ 1 / lgr_trap_rate,                            # the LGR sample rate
  #   str_detect(rel_group, fixed("LGR - HOR (AD-intact, CWT)) ~ 1 / (lgr_trap_rate * lgr_hor_subsample_rate),
  #   is.na(exp_rate) ~ ral_exp_rate,                                                    # if not lgr expansion rate, use bon rate
  #   TRUE ~ exp_rate
  # )) %>%
  group_by(rel_group,
           rel_year,
           ocean_age,
           node) %>%
  summarise(n_tags = sum(n_tags, na.rm = T),
            n_tags_exp = sum(n_tags_exp, na.rm = T),
            .groups = "drop")

```

```{r det-probs, results = "hide"}
# calculate detection probabilities
library(PITcleanr)
load(here("data/configuration_files/site_config_LGR_20250416.rda")) ; rm(flowlines, sr_site_pops)

node_paths = addParentChildNodes(parent_child, configuration) %>%
  buildNodeOrder()

# calculate node detection efficiencies for all spawn years
node_eff = sfsr_obs_yr %>%
  #group_by(spawn_year) %>%
  do(estNodeEff(capHist_proc = ., 
                node_order = node_paths)) %>%
  ungroup() %>%
  filter(node %in% c("SFG", "KRS_D", "KRS_U", "STR", "ESS_D", "ESS_U", "ZEN_D", "ZEN_D"))

```

```{r tag-exp-det-prob}
# expand estimates by site detection probabilities
exp_summ = tag_exp %>%
  filter(str_detect(rel_group, "McCall") | str_detect(rel_group, "LGR")) %>%
  # can't use tagging of ad-clipped adults at LGR i.e., exclude them... would be additive to juvenile releases
  #filter(rel_group != "LGR - HOR (AD-intact, CWT)") %>%
  left_join(node_eff %>%
              #filter(spawn_year == yr) %>%
              select(node, 
                     eff_est, 
                     eff_se)) %>%
  mutate(tot_est = round(n_tags_exp / eff_est)) %>%
  filter(!(node %in% c("KRS_U", "ESS_U", "ZEN_U"))) %>%
  # parse by adults vs. jacks
  mutate(age = case_when(
    #str_detect(rel_group, "LGR")        ~ "All",
    ocean_age == "2" | ocean_age == "3" ~ "Adults",
    ocean_age == "1"                    ~ "Jacks",
    TRUE ~ NA
  )) %>%
  group_by(rel_group,
           age,
           node,
           eff_est,
           eff_se) %>%
  summarize(n_tags = sum(n_tags),
            n_tags_exp = sum(n_tags_exp),
            tot_est = sum(tot_est),
            .groups = "drop") %>%
  mutate(est_tags = round(n_tags/eff_est)) %>%
  select(rel_group,
         age,
         node,
         n_tags,
         eff_est,
         eff_se,
         est_tags,
         n_tags_exp,
         tot_est) %>%
  arrange(rel_group,
          desc(node),
          age)

```

**Data Last Queried:** `r as.POSIXct(dt_tm, format = "%Y-%m-%d_%H-%M-%S") %>% format("%B %d, %Y %I:%M%p")`

To estimate the number of adults past each array (Table \@ref(tab:results-tbl)), we summarized the expanded counts, based on sort-by-code designation in Table \@ref(tab:tag-tbl), by release group and age (i.e., adults versus jacks) and further accounted for array detection efficiencies. Array detection efficiencies were estimated as the percent of tags detected at upstream array(s) that were detected at the given array.

PIT tag observations were queried from DART at the above date and time and summarized by release group, release year, and sort-by-code designation (i.e., run-at-large [RAL] versus return-to-river [RTR]) to determine the number of tags observed at the South Fork Guard Station (SFG) and Krassel (KRS) arrays on the South Fork Salmon River (Table \@ref(tab:tag-tbl)). The number of tags observed were then expanded based on expansion rates (inverse of mark rates). For in-season escapement estimates, we focus on natural-origin adults tagged at Lower Granite Dam (LGR) and juvenile hatchery releases to the South Fork Salmon River. For hatchery releases, expansion rates were derived from the **PIT Data** tab in "2025 PIT_Tag_Analysis_LowerGranite.xlsx". Natural-origin expansion rates were assumed to be `r 1/lgr_trap_rate` based on the set trapping and tagging rate at LGR (`r lgr_trap_rate * 100`%).


```{r results-tbl, tab.cap = "Estimated number of adults past each array by release group and age after expanding by mark rates and array detection efficiencies."}
results_tbl = exp_summ %>%
  filter(age == 'Adults') %>%
  mutate(node = factor(node, levels = c('SFG','KRS_D', 'ZEN_D', 'ESS_D'))) %>%
  select(-eff_se, -n_tags_exp) %>%
  mutate(eff_est = round(eff_est * 100)) %>%
  rename(`Release Group` = rel_group,
         Age = age,
         Array = node,
         `Tags Observed` = n_tags,
         `Array Det Eff` = eff_est,
         `Estimated Tags` = est_tags,
         `Est. Adults` = tot_est) %>%
  kable(booktabs = T,
        align = "ccccccc",
        label = NA) %>%
  kable_styling(full_width = F,
                position = "center",
                bootstrap_options = c("striped", "condensed"))
results_tbl

```


```{r run-timing-calcs}
sfsr_fish <- sfsr_obs %>%
  filter(mark_site %in% c('LGRLDR', 'MCCA')) %>%
  filter(rel_site %in% c('LGRLDR', 'KNOXB')) %>%
  mutate(julian = lubridate::mdy(paste0(format(min_det, format = '%m-%d'), '-2020'))) %>%
  # only interested in fish tagged at LGR or hatchery releases from MCCA sites in SF Salmon
  mutate(rel_group = case_when(
    rel_site == "KNOXB" & str_detect(flags, "AI")     ~ "McCall - Integrated",
    rel_site == "KNOXB" & str_detect(flags, "AD")     ~ "McCall - Segregated",
    rel_site == "LGRLDR" & mark_rear_type_name == "W" ~ "LGRLDR", # - NOR",
    rel_site == "LGRLDR" & str_detect(flags, "CW")    ~ "LGRLDR", # - HOR (AD-intact, CWT)",
    #rel_site == "KNOXB"  & mark_rear_type_name == "W" ~ "KNOXB - NOR",
    #rel_site == "SFSRKT" & mark_rear_type_name == "W" ~ "SFSRKT - NOR",
    TRUE ~ 'other'
  ))


obs <- sfsr_fish %>%
  filter(rel_group != 'other') %>%
  filter(node %in% c('SFG', 'KRS', 'KRS_D')) %>%
  mutate(node = ifelse(node == 'KRS', 'KRS_D', node)) %>%
  group_by(rel_group, node, tag_code) %>%
  slice_min(min_det)

# check run-timing by group - not a big difference
obs_sum <- obs %>%
  group_by(rel_group, node, spawn_year, julian) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  complete(rel_group, node, spawn_year, julian, fill = list(n=0)) %>%
  group_by(rel_group, node, spawn_year) %>%
  mutate(csum = cumsum(n),
         cdf = csum/sum(n)) %>%
  ungroup()

today = lubridate::mdy(paste0(format(Sys.Date(), format = '%m-%d'), '-2020'))

avg_run <- obs_sum %>%
  filter(spawn_year != yr) %>%
  group_by(rel_group, node, julian) %>%
  summarise(cdf = mean(cdf, na.rm = TRUE))

p <- obs_sum %>%
  filter(julian == today) %>%
  group_by(rel_group, node) %>%
  summarise(cdf = mean(cdf))

# avg_run %>%
#   filter(julian == today)
```


```{r run-timing, fig.align='center', fig.cap = "Run-timing of hatchery (segregated and integrated) and unclipped returns of LGRLDR released tags across South Fork Salmon River PIT-arrays. The grey lines indicate individual spawn year returns for 2010-2024, the dark blue line is the average cumulative proportion of returns, and the green line represents the current 2025 return."}

ggplot() +
    geom_line(data = obs_sum %>% filter(spawn_year != yr), aes(x = julian, y = cdf, group = spawn_year), colour = 'grey') +
    geom_line(data = avg_run, aes(x = julian, y = cdf), colour = 'navy', size = 1) +
    geom_line(data = obs_sum %>% filter(spawn_year == yr), aes(x = julian, y = cdf), colour = 'limegreen') +
    #geom_hline(xintercept = julian, colour = 'firebrick') +
    scale_x_date(date_breaks = '3 weeks', date_labels = "%b-%d") +
    facet_grid(rel_group ~ fct_rev(node)) +
  theme_bw() +
  labs(x = 'Date',
       y = 'CDF')
```

```{r run-timing-year, fig.align='center', fig.height=9, fig.cap = "Run-timing of hatchery (segregated and integrated) and unclipped returns of LGRLDR released tags across South Fork Salmon River PIT-arrays. The vertical red line indicates the last data download."}
ggplot() +
    stat_ecdf(data = obs, aes(x = julian, colour = rel_group), size = 1) +
    geom_vline(xintercept = today, colour = 'firebrick') +
    scale_x_date(date_breaks = '4 weeks', date_labels = "%b-%d") +
    facet_wrap(~ spawn_year, ncol = 2) +
  theme_bw() +
  theme(legend.position = 'top') +
  labs(x = 'Date',
       y = 'CDF',
       colour = '')
```

```{r tag-tbl, tab.cap = "PIT tags detected at the SFG and KRS sites by release group, release year, and sort-by-code designation including expansion rates and an estimate of the number of adults at each array, **not accounting for site detection probability.**"}
tag_tbl = tag_df %>%
  #filter(rel_group != "LGR - HOR (AD-intact, CWT)") %>%
  filter(node %in% c("SFG", "KRS_D", "ZEN_D", "ESS_D")) %>%
  select(rel_group,
         rel_year,
         sby_c,
         node,
         n_tags,
         exp_rate,
         n_tags_exp) %>%
  arrange(rel_group,
          desc(node)) %>%
  rename(`Release Group` = rel_group,
         `Release Year` = rel_year,
         `Sort-By-Code` = sby_c,
         Array = node,
         `Tags Observed` = n_tags,
         `Expansion Rate` = exp_rate,
         `Mark Rate Expanded` = n_tags_exp) %>%
  kable(booktabs = T,
        align = "ccccccc",
        label = NA) %>%
  kable_styling(full_width = F,
                position = "center",
                bootstrap_options = c("striped", "condensed"))
tag_tbl

```